### 2022-06-05 manuscript.Rmd
Update manuscript.Rmd
Added to Starting Values
Added Assessing Estimation Results

The EM algorithm needs starting values for each parameter it wants to estimate. 
If it were provided a single starting value for each parameter, 
the EM optimizer would be at risk of finding a local optimum instead of the global optimum. 
In other words, it would converge on a suboptimal solution resulting in biased parameter and model fit estimates. 
To prevent this, we should give the algorithm a sufficiently large set of random starting values for each parameter in order to maximize our chances of finding the deepest valley that is the global optimum (McLachlan & Peel, 2000). Furthermore, we should ensure the number of initial stage iterations is large for optimization to be successful (Geiser, 2013).

@@

Each set of starting values will result in a model with a log likelihood value. Ideally, the largest log likelihood value will be replicated a large number of times (each time is generated by a different starting value set). Successful replications of the largest log likelihood value across different starting values instill confidence that the estimation was successful. The model with the highest replicated likelihood will be chosen as our best-fit. If the largest log likelihood value is not replicated many times, the model may be unidentified (Geiser, 2013).

[to add: Nylund-Gibson & Choi  p.22. about 3-10% and Geiser p. 237 -240]

>>>>

The EM algorithm needs starting values for each parameter it wants to estimate. 
If it were provided a single starting value for each parameter, 
the EM optimizer would be at risk of finding a local optimum instead of the global optimum. 
In other words, it would converge on a suboptimal solution resulting in biased parameter and model fit estimates. 
To prevent this, we should give the algorithm a sufficiently large set of random starting values for each parameter in order to maximize our chances of finding the deepest valley that is the global optimum (McLachlan & Peel, 2000). Furthermore, we should ensure the number of initial stage iterations is large for optimization to be successful (Geiser, 2012).

@@

Each set of starting values will result in a model with a log likelihood value. Ideally, the largest log likelihood value will be replicated a large number of times with each log likelihood value generated by a different starting value set. Successful replications of the largest log likelihood value across different starting values instill confidence that the estimation was successful. The model with the highest replicated likelihood will be chosen as our best-fit. If the largest log likelihood value is not replicated many times, the model may be unidentified (Geiser, 2012). Nylund-Gibson and Choi (2018) recommend that the highest log likelihood should be replicated in at least 3-10% of the models initialized with different starting values in the first step of the optimization.

Another consideration has to do with conditional item probabilities which indicate the probability of a positive score on an item given that the person belongs to a particular latent class. If many conditional item probabilities in estimation are found to be extreme (as in exactly 0 or 1), we are observing boundary parameter estimates. This can be a sign of an invalid solution, warn us that too many classes were extracted, or indicate a local optimum. Boundary parameter estimates are more likely to happen when working with sparse data and should be interpreted with caution (Geiser, 2012).
